---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Impacts of Dam Removal on the Neuse River"
subtitle: "https://github.com/lydiecos/Neuse-Dam-Removal"
author: "Lydie Costes"
fontsize: 12pt
mainfont: Times New Roman

---

\newpage

```{r setup, include=FALSE}
# Get working directory
getwd()

# Load packages
library(tidyverse)
library(lubridate)
library(dataRetrieval)

# Set ggplot theme
theme_set(theme_bw())

# Load datasets
# Download discharge data
NeuseFlow <- readNWISdv(siteNumbers = "02089500",
                    parameterCd = "00060",
                     startDate = "",
                     endDate = "")

# Download water quality data
NeuseWQ  <- readWQPqw(siteNumbers = "USGS-02089500",
                      parameterCd = c("00915", # calcium, filtered mg/L
                                      "00925", # magnesium, filtered mg/L
                                      "00930", # sodium,filtered mg/L
                                      "90095", # Specific conductance, uS/cm
                                      "31625", # Fecal coliform, CFU/100 mL
                                      "31673"), # Fecal streptocci, CFU/100 mL
                      startDate = "",
                      endDate = )

```


# Rationale and Research Questions



\newpage

# Dataset Information

```{r wrangling}
# Change column names
names(NeuseFlow)[4:5] <- c("Discharge", "Approval.Code")
# Check Flow data formats
str(NeuseFlow)
# Check Flow data summary
summary(NeuseFlow)
# View flow over time
ggplot(NeuseFlow, aes(x = Date, y = Discharge)) +
  geom_line() 

# Convert WQ to wider dataframe with characteristics of interest
NeuseWQ_processed <- NeuseWQ %>%
  select(MonitoringLocationIdentifier, ActivityStartDate, HydrologicCondition, 
         CharacteristicName, ResultMeasureValue) %>%
  mutate(Variable = case_when(CharacteristicName == "Fecal Coliform" ~ "FecalColiform_cfu.100mL", 
                              CharacteristicName == "Fecal Streptococcus Group Bacteria" ~ "FecalStrep_cfu.100mL",
                              CharacteristicName == "Specific conductance" ~ "SpC_uS.cm",
                              CharacteristicName == "Calcium" ~ "Calcium_mg.L",
                              CharacteristicName == "Magnesium" ~ "Magnesium_mg.L",
                              CharacteristicName == "Sodium" ~ "Sodium_mg.L")) %>%
  select(-CharacteristicName) %>%
  pivot_wider(names_from = "Variable", values_from = "ResultMeasureValue")

# Rename the date column
names(NeuseWQ_processed)[2] <- "Date"
# Check WQ data formats
str(NeuseWQ_processed)
# Check WQ data summary
summary(NeuseWQ_processed)

# The two fecal variables have too many NAs; remove them
NeuseWQ_processed <- NeuseWQ_processed %>%
  select(-FecalStrep_cfu.100mL, -FecalColiform_cfu.100mL)

# View Conductivity over time
ggplot(NeuseWQ_processed, aes(x = Date, y = SpC_uS.cm)) +
  geom_point()
# View Sodium over time
ggplot(NeuseWQ_processed, aes(x = Date, y = Sodium_mg.L)) +
  geom_point()
# View Calcium over time
ggplot(NeuseWQ_processed, aes(x = Date, y = Calcium_mg.L)) +
  geom_point()
# View Magnesium over time
ggplot(NeuseWQ_processed, aes(x = Date, y = Magnesium_mg.L)) +
  geom_point()
```

\newpage

# Exploratory Analysis 



\newpage

# Analysis



## Question 1: Has there been an increase in release of various elements over time? Have levels steadily increased since dam removal, or did they spike and then decline, as we might hope to see?

## Question 2: Have discharge levels become more extreme since dam removal? (recognizing that separating the effects of dam removal versus climate change may be challenging)




\newpage

# Summary and Conclusions


\newpage

# References
<add references here if relevant, otherwise delete this section> 
